The Multi-Stencil Language: orchestrating stencils with a mesh-agnostic DSL

### 2016-05-30 ARTICLES/IJPP/article/introduction.tex

One possible runtime execution model for HPC is to propose dynamic scheduling of task graphs combined to message passing models~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} (to be able to use more than one machine). Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming. At a highest abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models. However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture. The current easiest (with the higher abstraction level), but still efficient, programming model for end-users is Domain Specific Language (DSL). Such a language is specific to the end-user domain and proposes a grammar which is easy to understand. The DSL compiler is able, because of the specific knowledge onto the domain targetted, to automatically apply parallelization and specific optimizations to produce a high performance back-end code. Thus, a DSL is able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.

>>>>

One possible runtime execution model for HPC is to propose dynamic scheduling of task graphs combined to message passing models~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} (to be able to use more than one machine). Those models increase HPC code portability and reach an interesting performance onto heterogeneous architectures, which is interesting to reach exascale programming. At a higher abstraction level, general purpose parallel languages, such as OpenMP~\cite{660313} and OpenCL~\cite{Stone:2010:OPP:622179.1803953} follow the direction of task graph scheduling proposed by those execution models. However, for non-experts end-users, general purpose languages still are difficult to use and to tune for a given application onto a given architecture. The current easiest (with the higher abstraction level), but still efficient, programming model for end-users is Domain Specific Language (DSL). Such a language is specific to the end-user domain and proposes a grammar which is easy to understand. The DSL compiler is able, because of the specific knowledge onto the domain targetted, to automatically apply parallelization and specific optimizations to produce a high performance back-end code. Thus, a DSL is able to split end-user concerns from HPC concerns which is relevant to reach exascale programming models.
